Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-29 ltx/paper.tex
more edits. new figures. re-do figures for pets2000 and pets2001 datasets

% \begin{figure}[h]
%   \centering
%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
%   \label{fig:baselinedata}
% \end{figure}

@@

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
P(z_{k,t,m} | state) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

@@

\subsubsection{Initializing the Sampler}
Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized. While the state of the sampler can theoretically be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.

@@

The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA). These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.

The following terms are used in the definitions of the performance metrics:
\begin{itemize}
\renewcommand{\labelitemi}{$\bullet$}
\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
\end{itemize}

@@

\subsubsection{Tracking by Appearance}

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_cross.pdf}}
  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_bounce.pdf}}
  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth4.pdf}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.}
  \label{fig:synth_one_plot}
\end{figure}

\subsubsection{Tracking though Occlusion and Appearance Shift}

@@

\begin{figure}[h]
        \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
        \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown. Both the sample and data are overlaid on three images from synthetic experiment. The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.  \willie{will show assignments by marker type in addition to color}}
\end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three points in time, where the data is colored by assignment. \willie{also show pets 2001 result and per-frame overlays for both}.

\begin{figure}[h]
        \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
        \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
\end{figure}

\begin{figure}[h]
	\centering
	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
	\label{fig:pets2000_results}
\end{figure}

\begin{figure}
  \centering
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}             
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_back.pdf}}
  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b). Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
  \label{fig:benchmark_results}
\end{figure}

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f1.pdf}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f3.pdf}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f5.pdf}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f7.pdf}}
  \caption{PETS2001 results.  \willie{add tails. different color ovals for different objects.}}
  \label{fig:pets2001_overlay}
\end{figure}

@@

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames.} \willie{These are old images!... haven't replaced yet}}
  \label{fig:pets2001_imgs}
\end{figure}

@@

The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}). Figure \willie{provide figure showing PETS2009 results} shows a sample from the posterior of the cluster parameters overlayed on the data, and video frames at three points in time, where the data is colored by assignment
\willie{show SFDA and ATA vs cov mat conf for this sample that I show figures for.}

>>>>

% frank says the following worm figures are a no go
% \begin{figure}[h]
%   \centering
%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
%   \label{fig:baselinedata}
% \end{figure}

@@

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
P(z_{k,t,m} | state) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

@@

\subsubsection{Initializing the Sampler}
Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized. While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.

@@

The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA). These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking. The following are used in the definitions of the performance metrics:
\begin{itemize}
\renewcommand{\labelitemi}{$\bullet$}
\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
\vspace{2mm}
\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
\vspace{2mm}
\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
\vspace{2mm}
\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
\vspace{2mm}
\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
\vspace{2mm}
\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
\end{itemize}

@@

% \subsubsection{Tracking by Appearance}

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_cross.pdf}} \hspace{0.5mm}
  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_bounce.pdf}} \hspace{0.5mm}
  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth4.pdf}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.}
  \label{fig:synth_one_plot}
\end{figure}

% \subsubsection{Tracking though Occlusion and Appearance Shift}

@@

% \begin{figure}[h]
%         \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
%         \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown. Both the sample and data are overlaid on three images from synthetic experiment. The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.  \willie{will show assignments by marker type in addition to color}}
% \end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figure, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

\begin{figure}[h]
  \centering
  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}}
  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b). Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
  \label{fig:benchmark_results_1}
\end{figure}

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}}
  \caption{Four frames from the PETS2000 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
  \label{fig:pets2001_imgs}
\end{figure}

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}}
  \caption{Four frames from the PETS2001 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
  \label{fig:pets2001_overlay}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
    \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown. In each plot, the sample and data are overlaid on three images from the experiment.}
\end{figure}

% old figure ideas:

% \begin{figure}[h]
%         \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
%         \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
% \end{figure}

% \begin{figure}[h]
% 	\centering
% 	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
% 	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
% 	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
% 	\label{fig:pets2000_results}
% \end{figure}

@@

The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}). In Figure \ref{fig:pets2009_results}, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.

\begin{figure}[h]
  \centering
  % \subfloat[]{\label{fig:pets2009_results} \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}}
  \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}
  \caption{Results on frames 1-50 of the PETS2009 dataset, showing a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
  \label{fig:pets2009_results}
  % \label{fig:benchmark_results_2}
\end{figure}
